---
title: "p8105_hw6_ml4198"
author: "Matthew Lawlor"
date: "12/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
```

## Problem 1

### Read and tidy

```{r}
homicide_df = read.csv("./data/homicide-data.csv") %>% 
  janitor::clean_names() %>% 
  filter(city != "Dallas", city != "Phoenix", city != "Kansas City", city != "Tulsa", victim_race %in% c("White", "Black")) %>% 
  mutate(
    city_state = paste(city, state, sep = ", "),
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")
  ) %>% 
  select(resolved, city_state, victim_age, victim_sex, victim_race)
```

### Fit logistic regression model

```{r}
fit_baltimore =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial())
```

Now tidy

```{r}
fit_baltimore %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>% 
  select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3)
```

Adjusted OR for case resolved in black v white victims is 0.431.

### Regress all cities

```{r}
fit_all =
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    model = map(.x = data, ~glm(resolved ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(model, broom::tidy)
  ) %>% 
  unnest(results) %>%
  filter(term == "victim_raceBlack") %>% 
  mutate(
    OR = exp(estimate),
    ci_lower = exp(estimate - std.error),
    ci_upper = exp(estimate + std.error)
  ) %>%  
  select(city_state, OR, ci_lower, ci_upper)
```

then plot

```{r}
fit_all %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = OR, y = city_state)) +
  geom_vline(aes(xintercept = 1)) +
  geom_errorbarh(aes(xmax = ci_upper, xmin = ci_lower), size = .5, height = .2, color = "gray50") +
  geom_point(color = "orange")
```

This is a plot of OR +/- CI for resolved cases for black vs white victims. There is variation across cities for odds of resolved case regressing by race. Unfortunately, many cities lie well below OR 1 suggesting homicides with black victims are less likely to be solved.
